### 2021-06-28 article/article.Rmd
made outline of sections, looking to make planned fig1 next time

---
title: "Osa"
output: html_notebook
---

# Introduction

1.  importance of reforestation

2.  patterns of reforestation

    1.  general

    2.  tropical

3.  things that affect reforestation

    1.  edge effects

    2.  plantation structure?

4.  [for NGO--start biomass census]

    1.  Does leaving the plantation structure help reforestation?

        1.  shade

        2.  edge effects (i.e. buffer)

# Results

## Shade / Canopy

1.  shade x distance + timber sp

## Edge effects

1.  biomass x distance + timber sp

    -   can break down biomass by component

2.  diversity x distance [+ 1 timber sp]

    -   can break down by diversity index

# Discussion

1.  assess hypotheses
2.  infer causes of patterns - plantation structure, old legacy trees, or species traits?
3.  compare inferences to other studies
4.  recommend future conservation approaches, ecological studies

# Methods

## Study site

-   osa peninsula (GPS)
-   wet lowland rainforest
-   greg gund conservation center, osa conservation

## Plot design

-   QGIS/ArcGIS 50m bands 0-300m from forest
-   30 plots, random within each band, allocated by area: 11, 8, 5, 3, 2, 1
-   20x20m

## Data collection

### Canopy light and slope

-   densiometer in 4 directions at center of plot
-   slope?

### Tree biomass

-   DBH, H using rangefinder (BUSHNELL)

-   chave 2014, updated to chave 2017, little difference in result

-   unique individual tag/plate \#s

-   species IDs

    -   local knowledge of common names

    -   taxonomy matches using local field guides

-   wood density using Agroforestry Wood Database in 2013, updated using BIOMASS package

## Statistical analysis

-   code available (somewhere..., DRYAD, GitHub)

-   R v\#\#\#

    -   BIOMASS package v\#\#\#

        -   taxa (family, maybe genus) updated from raw IDs using field guides

        -   matched wood densities from package database

    -   tidyverse v\#\#\# = summary stats, plotting

    -   vegan v\#\#\# = community analysis

# Acknowledgements

-   osa conservation - max villalobos, [dan student harvard alum], agustín (& annia), manuel director
-   lehigh group - don, emma, ?, ?
-   brandeis WOW funding