### 2022-03-04 talk/talk.Rmd

---
title: 
  "Natural regeneration of a tropical timber plantation"
author: 
  "Nicholas Medina"
output: 
  # slidy_presentation
  # oislides_presentation
  # powerpoint_presentation
  html_notebook
  # html_document
  # pdf_document #no
---

```{r setup, echo = F, include = F}
knitr::opts_chunk$set(
  echo = F,
  quiet = T
)
library(
  here
)
i_am(
  "talk/talk.Rmd"
)
# #fixRender
# library(
#   httpuv
# )
#embed_file()
# library(
#   xfun
# )
# xfun::pkg_load2(c("htmltools", "mime"))
```

Outline
===

- Introduction
- Methods
- Results
- Discussion

I--Global forests are regenerating, but how?
===

- ~Half of forests are tropical, and over half of all forests globally are secondary *(FAO 2020)*

- Services like timber and biodiversity depend on natural and managed regeneration patterns

- Regeneration dynamics should be better understood

I--Regeneration is quick and inconsistent
===

- Biomass recovers 90% in 66 yrs *(Poorter et al 2016 PNAS)*

- Taxonomic richness recovers quickly ()

- Community composition may take centuries to recover, if at all *(Norder et al 2010 PNAS)*

I--Edges shape regeneration
===

- 70% forests are <1 km from a habitat edge *(Haddad et al 2015 Science)*

- Edges effects tend to reduce biodiversity by 13 - 75% via ... *(")*

    - stronger in small and isolated fragments

I--Edge effects may differ in wet tropical plantations
===

- 7% of global forests are planted and planted area has increased over last 30 yrs *(FAO 2020)*

    - Often dense with fast-growing, shade-intolerant taxa

- Wet tropical forest regeneration tends to be light-limited, based on tree functional traits *(Poorter et al 2021)*

- Existing plantation demography and shade may skew/buffer edge effects toward net positive instead of negative

I--This study
===

- Question -- How is wet tropical secondary forest regeneration affected by former plantation status and edges with road and primary forest?

- Hypothesis --

    - Former plantation status facilitates succession by allowing shade-intolerant taxa to grow better due to __denser canopies__

    - Forest edges promotes regeneration by faciltating __dispersal__ of shade-tolerant taxa

- Predictions -- Distance away from primary forest edge is associated with --

    - lower stand biomass

    - lower tree diversity

    - lower shade-tolerant taxa

II--Study site
===

- Wet tropical rainforest in Osa Peninsula, SW Costa Rica

- 20 ha plantation of dry forest timber, abandoned in ~2003 into NRDC preserve

- Area surrounded by primary forest on two sides and road on third

```{r site}
# nature_img.png
```

II--Census design
===

- Random stratified forest biomass inventory sampling design --

    - divided 300 m range into six 50 m strata, from forest to road edge, using QGIS and ArcGIS

    - 30, 21 x 21 m plots randomly distributed, but quantity roughly proportional to stratum area

    - All plots covered 1 ha total, 5% of entire plantation

![](../figs/map.png)

II--Census measurements
===

- Stems ≥10 cm were measured for DBH and height using rangefinder *(Bushnell, Forestry Suppliers, Inc)*

- Taxa were ID'd with help from local field guides

- Plot canopy closure was measured with densiometer

II--Statistical analysis
===

- For all variables--

    - Trend lines follow median values of all plots in distance stratum

- Traits obtained from literature and compared to TRY database *(Kattge et al 2020)*

- Computations done with R `getRversion()` depending on *BIOMASS* `packageVersion("BIOMASS")`, 
*vegan* `packageVersion("vegan")`, and 
*tidyverse* `packageVersion("tidyverse")` packages

III--Stand regeneration
===

- Further from primary forest edge--

    - Biomass and wood density tended to increase

    - Stem density decreased then increased

    - Neither canopy light nor height (not shown) tended to change

- Distance from forest edge tended to explain variation--

    - 10% in wood density

    - ~20% in stem density

```{r fig1}
# source(
#   here(
#     "figs/fig1.R"
#   )
# )
# fig1
```

![](../figs/fig1.png)

III--Composition regeneration
===

- Further from forest edge--

    - Richness tended to increase

    - Diversity decreased steeply

    - Composition varied notably, toward PC1

    - Of abundant taxa biomass--

        - Vochysia tended to decrease

        - Ficus decreased

```{r fig2}
# source(
#   here(
#     "figs/fig2.R"
#   )
# )
```

![](../figs/fig2.png)

III--Functional regeneration
===

-

-

-

```{r fig3}
# source(
#   here(
#     "figs/fig3.R"
#   )
# )
# fig3
```

![](../figs/fig3.png)

IV--Hypothesis check
===

- Stand biomass is more resilient than biodiversity

    - Further from forest--

        - No trends in biomass

        - Diversity decreased

- Forest edge associated with tree diversity

IV--Processes underlying patterns
===

- not shade...

- dispersal...

-

IV--Compare to similar studies
===

- aligns well with major recent papers

-

-

IV--Future studies
===

- Compare wet tropical plantations at different stages of regeneration

- Compare forests with different initial composition but in similar region (i.e. realized vs. fundamental niches)

- Study ecological dynamics under various management conditions

<!-- For more details on authoring R presentations please visit <https://support.rstudio.com/hc/en-us/articles/200486468>. -->