### 2022-06-18 article/disc.Rmd
minors, prep disc

---
title: "Discussion"
output: 
  # html_notebook
  bookdown::html_document2:
    # number_sections: FALSE
bibliography: ../../library.bib
link-citations: TRUE
---

# outline{-}
1.  assess hypotheses
2.  infer causes of patterns - plantation structure, old legacy trees, or species traits?
3.  compare inferences to other studies
4.  recommend future conservation approaches, ecological studies

# draft{-}

This study found significant edge effects on tree diversity and community composition, but not aboveground biomass...functional diversity?
These results were shown in Figs \@ref(fig:biomass), \@ref(fig:taxa), and \@ref(fig:traits)...
It was initially hypothesized that community regeneration would vary with distance to edge, favoring more shade-tolerant taxa, due to light availability and dispersal potential -- evidence from this study's results supported edge effects on community composition and biomass of taxa associated with both early- and late-successional stages, but without detectable roles for light availability or general dispersal mode.
...

Aboveground biomass represents aboveground carbon stores... is a measure of ..., and responds variably to secondary succession and edge effects

# attic{-}

Under more sustainable forestry that maintains multiple functions, forest stand regeneration is likely more predictable, generally mirroring (i.e. near 1:1 replacement) the targeted tree cutting that is done using information collected about the ecological profiles and ages (when possible) of trees (Duah-Gyamfi et al 2014, Edwards et al 2014).

In contrast, under intensive silviculture that prioritizes the growth of select tree crop species, forest stands can show high wood production (Ferez et al 2015) while also usually showing negative trade-offs with other services, like that with biodiversity given reliance on monoculture planting designs.

In some cases where management decisions are made primarily based on regional economic but not local ecological information, due to limited access to funded scientific reporting (Cook et al 2012), forest stands can perform very poorly all-around.

After this more extreme type of management, it remains uncertain how natural regeneration in secondary forests will proceed, how local ecosystem services will be restored, and how restoration could be managed locally (Laurance et al 2012), e.g. based on surrounding forest proximity.

The persistence of early-successional tree communities has been found to lower the standing biomass, likely due to lower wood density *[@césar21]*, and could therefore require a switch in focal timber species *[@karsten14]*.