### 2022-07-11 article/disc.Rmd

---
title: "Discussion"
output: 
  # html_notebook
  bookdown::html_document2:
    # number_sections: FALSE
bibliography: ../../library.bib
link-citations: TRUE
---

<!-- # outline{-} -->
<!-- 1.  assess hypotheses -->
<!-- 2.  infer causes of patterns - plantation structure, old legacy trees, or species traits? -->
<!-- 3.  compare inferences to other studies -->
<!-- 4.  recommend future conservation approaches, ecological studies -->

<!-- # draft{-} -->

This study found significant edge effects on secondary forest stand wood and stem density, tree diversity and community composition, and some functional traits, yet not on overall aboveground biomass.
<!-- These results were shown in Figs \@ref(fig:biomass), \@ref(fig:taxa), and \@ref(fig:traits)... -->
It was initially hypothesized that community composition would vary with distance to edge, favoring more shade-tolerant taxa, due to light availability and dispersal potential, and this evidence supported edge effects on community composition and biomass of taxa associated with both early- and late-successional stages, but without detectable roles for light availability or general dispersal mode.
Accordingly this suggests that edge effects do significantly change humid tropical forest community composition, but primarily due to factors other than light competition or dispersal limitation, as well as that aboveground biomass and even low taxonomic richness is resilient after a decade of regeneration.

Aboveground biomass storage is a key global ecosystem function and service, but edge effects tend to release about one-third of carbon lost from tropical deforestation *[@brinck17]*, with more expected in the future *[@mitchard18]*.
<!-- responds variably to secondary succession and edge effects ...   -->
Management has been found to explain most of biomass variation among tropical forest fragments, and wood density ~10% biomass variation *[@pyles22]*.
This study adds that management via edge effects can also mediate ~10% of wood density variation without affecting local biomass, although median stem wood density can describe regional spatial biomass patterns *[@baker04]*.
Overall, wood density and richness values may have been pulled down by early-successional clonal trees closest to the primary forest, which tend to have lower wood densities, lowering carbon *[@gonzalezj98; @pyles22]*, and increasing nearby stem density.
Lower wood density can also reflect more dynamic forest stands *[@malhi06]* with higher tree turnover, a consequence of forest fragmentation *[@nascimento04a]* together with other global change drivers *[@laurance14]*, as well as may reflect possible differences in soil fertility *[@malhi04]*, all of which represent possible future research directions, beyond strictly neutral dynamics *[@terborgh95]*.
Furthermore, light availability at the plot level may not affect biomass storage, but instead individual light gaps may be more important for biomass dynamics *[@chazdon84]*.
    <!-- - although canopy light may still affect biomass generally in this case, without edge effects (check direct light-biomass correlation graph? -- no sig correlation) -->

Edge effects have been shown to lower both forest biomass and also diversity, and this study adds that the decline can be rapid and non-linear across edge gradients, along with supporting the likelihood of lasting changes to community composition.
The rapid diversity loss across the edge gradient may be in part due to finer variations in dispersal patterns among animal taxa, rather than overall dispersal mode more broadly.
Similarly, stochastic factors like priority effects after dispersal may also affect offspring establishment and re-growth.
While the biomass of the most abundant taxon *Vochysia* was resilient to edge effects, that of the key taxon for frugivores like birds and bats *Ficus* was significantly lowered across the forest edge, pointing to potentially different responses of biomass and food web associated ecosystem services.

<!-- # draft attic{-} -->

<!-- This study supports previous syntheses suggesting that forest biomass and taxonomic richness regenerate more quickly / are more resilient than diversity and community composition *[@poorter21?, other?]*. -->
<!-- This may be due to priority effects on successional trajectories acting on/at the stem/population level *[]*, variable turnover and reproduction rates among tree taxa *[]*, and/or differential dispersal and establishment success based on local [dispersive] animal community *[]*. ... -->
<!-- - mention more amundant genera patterns w/ *Vochysia* and *Ficus* as a key fruit-bearing taxon *[@tropicalbat/bird]*? -->

<!-- There is increasing interest in functional trait ecology of tropical trees *[@umaña##]*. -->
<!-- This focuses on a coarse functional association with successional stage, ... -->
<!-- but also includes patterns with wood density [underlying biomass estmations]... -->
<!-- - fig3 -- edge effect on succession generalists, but not specialists -->
<!-- The persistence of early-successional tree communities has been found to lower the standing biomass, likely due to lower wood density *[@césar21]*, and could therefore require a switch in focal timber species *[@karsten14]*. -->

<!-- # attic{-} -->

<!-- Under more sustainable forestry that maintains multiple functions, forest stand regeneration is likely more predictable, generally mirroring (i.e. near 1:1 replacement) the targeted tree cutting that is done using information collected about the ecological profiles and ages (when possible) of trees (Duah-Gyamfi et al 2014, Edwards et al 2014). -->

<!-- In contrast, under intensive silviculture that prioritizes the growth of select tree crop species, forest stands can show high wood production (Ferez et al 2015) while also usually showing negative trade-offs with other services, like that with biodiversity given reliance on monoculture planting designs. -->

<!-- In some cases where management decisions are made primarily based on regional economic but not local ecological information, due to limited access to funded scientific reporting (Cook et al 2012), forest stands can perform very poorly all-around. -->

<!-- After this more extreme type of management, it remains uncertain how natural regeneration in secondary forests will proceed, how local ecosystem services will be restored, and how restoration could be managed locally (Laurance et al 2012), e.g. based on surrounding forest proximity. -->