### 2022-07-11 article/merge.Rmd

---
title: "Habitat edges affect tree diversity more than biomass regeneration in a wet neotropical secondary forest"
author: 
  - Nicholas Medina #^a^*
  # - Max Villalobos? ^b^
  # - Andrea Johnson? ^b^
date: 
# scriptsize\
  Ecology and Evolutionary Biology, University of Michigan, 1105 N University Ave, Ann Arbor, MI, USA 48109
  # ^a^ Biology and Environmental Studies, Brandeis University, 415 South St, Waltham, MA, USA 02453; 
  # ^b^ Land Stewardship, Osa Conservation, Puerto Jiménez, Golfito, Costa Rica
# author: 
#   - name: "Nicholas Medina"
#     email: nmedina@umich.edu
#     affiliation: Brandeis University
#     orcid_id: 0000-0001-5465-3988
# address:
#   - code: Brandeis University
#     address: Biology and Environmental Studies, 415 South St, Waltham, MA 02453
# keywords: [1, 2, 3]

output: 
  # html_notebook
  # bookdown::html_document2:
  #   # df_print: paged
  #   global_numbering: TRUE
  #   toc: FALSE
  bookdown::pdf_document2:
    # number_sections: FALSE
    global_numbering: TRUE
    keep_tex: TRUE
    toc: FALSE
bibliography: 
  - ../../library.bib
  - ../../grateful-refs.bib
  - ../../software.bib
link-citations: TRUE
csl: ../../styles-master/elsevier-harvard.csl

fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespace
# abstract
---

<!-- journals: J [Appl] Veg Sci, Basic Appl Ecol (3.5), Biotropica (2.5), Eco Evo -->

\
\

**Keywords**: secondary forest; regeneration; timber plantation associated biodiversity; edge effect; aboveground biomass; shade

\
\

***Corresponding author**: Nicholas Medina, nmedina@umich.edu

\newpage

# Abstract{-}

```{r child='abs.Rmd'}
```

\newpage

# Introduction

```{r child="intro.Rmd"}
```

# Methods

```{r child="methods.Rmd"}
```

# Results

```{r child="results.Rmd"}
```

# Discussion

```{r child="disc.Rmd"}
```

\newpage

# Funding{-}

This study was funded by Brandeis University Hiatt Career Center World of Work Fellowship.

# Declaration of interests{-}

Authors declare no conflicts of interest.

# Acknowledgements{-}

Thanks to Professor Dan Perlman for initial networking; concurrent Osa Conservation managers Max Villalobos and Andrea Johnson for project guidance and supervision; Lehigh University Professor Don Morris and interns Emma Capetz and Erin Lau along with field station staff Agustín Mendoza for direct field assistance; concurrent Osa Conservation general manager Manuel Ramirez and Piro Biological Station staff Juan Carlos Cruz Díaz, Annia Barrantes, Larry Villalobos, Hansel Vargas, and visitors for field support; peer James Fifer for discussion of early draft; and Luisa Valdez for translation assistance.

# Author contributions{-}

<!-- MV and AJ contributed to study conceptualization, administration, and supervision; NM and AJ contributed to funding acquisition; NM contributed to data collection; NM and MV contributed to data analysis;  -->
NM wrote initial draft and later revisions.

# Data statement{-}

Code stored at [github.com/nmedina17/osa](github.com/nmedin17/osa).

\newpage

# References